Computing in the Field of Rationals

### 2012-06-21 rational_ext.tex
_

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{\asterix}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o+ = \multimap^{+}
%! -o* = \multimap^{\times}
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! Pi = \Pi
%! Pi0 = \Pi^{o}
%! PiEE* = \Pi^{\eta\epsilon}_{*}
%! PiEE+ = \Pi^{\eta\epsilon}_{+}
%! PiEE = \Pi^{\eta\epsilon}
%! CatSet = \textbf{Set}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

To summarize, the syntax of {{Pi}} is given as follows.

@@

%%%%%%%%%%%%%%%
\subsection{Graphical Language}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Semantics}

@@

%subcode{proof} include main
%@@ f v1 |--> v2
%@ g v2 |--> v3
%@ h v3 |--> v4
%@@ g (;) h  v2 |--> v4
%@@@ f (;) (g (;) h) v1 |--> v4

and

%subcode{proof} include main
%@ f v1 |--> v2
%@ g v2 |--> v3
%@@ f (;) g v1 |--> v3
%@@ h v3 |--> v4
%@@@ (f (;) g) (;) h v1 |--> v4

@@

\begin{lemma}[Dagger]
  {{Pi}} is a dagger category, where every morphism {{f : b1<->b2}}
  has the adjoint {{f^{dagger}:b2 <->b1}}. The following properties
  hold (where {{g : b2 <-> b3}}): 
  \begin{enumerate}
  \item {{id^{dagger} = id : b <-> b}}
  \item {{(f (;) g)^{dagger} = g^{dagger} (;) f^{dagger}: b3 <-> b1}}
  \item {{f^{dagger dagger} = f : b1 <-> b2}}
  \end{enumerate}
\end{lemma}

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assoc+}}
    \item {{0 + b <-> b}} given by {{zeroe}}
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1<->b1}(+)id_{b2<->b2} = id_{b1+b2<->b1+b2} }}
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}
      \end{enumerate}
    \item We need to check the naturality of {{assoc+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assoc+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assoc+}}
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the ``pentagon'' and ``triangle'' axioms (see Sec 3.1
      \cite{springerlink:10.1007/978-3-642-12821-94})
    \end{enumerate}

@@

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon axioms''.
    \item The braiding is self inverse
      {{swap+_{b1+b2}=(swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}

>>>>

%subcode-inline{bnf-inline} regex \{\{(((\}[^\}])|[^\}])*)\}\} name main include langRev
%! Gx = \Gamma^{\times}
%! G = \Gamma
%! [] = \Box
%! |-->* = \mapsto^{\asterix}
%! |-->> = \mapsto_{\ggg}
%! |--> = \mapsto
%! <--| = \mapsfrom
%! |- = \vdash
%! <><> = \approx
%! ==> = \Longrightarrow
%! <== = \Longleftarrow
%! <=> = \Longleftrightarrow
%! <-> = \leftrightarrow
%! ~> = \leadsto
%! -o+ = \multimap^{+}
%! -o* = \multimap^{\times}
%! -o = \multimap
%! ::= = &::=&
%! /= = \neq
%! @@ = \mu
%! [^ = \lceil
%! ^] = \rceil
%! forall = \forall
%! exists = \exists
%! empty = \epsilon
%! Pi = \Pi
%! Pi0 = \Pi^{o}
%! PiEE* = \Pi^{\eta\epsilon}_{*}
%! PiEE+ = \Pi^{\eta\epsilon}_{+}
%! PiEE = \Pi^{\eta\epsilon}
%! CatSet = \textbf{Set}
%! theseus = Theseus
%! sqrt(x) = \sqrt{#x}
%! surd(p,x) = \sqrt[#p]{#x}
%! inv(x) = \frac{1}{#x}
%! frac(x,y) = \frac{#x}{#y}
%! * = \times

@@

\noindent
To summarize, the syntax of {{Pi}} is given as follows.

@@

\begin{definition}[Size of a type]
\label{def:size}
The size of a type {{b}} is a number defined to be:
\vspace{-20pt}
\begin{multicols}{2}
%subcode{opsem} include main
% [^ b1 + b2 ^] '= [^ b1 ^] + [^ b2 ^]
% [^ b1 * b2 ^] '= [^ b1 ^] * [^ b2 ^]

%subcode{opsem} include main
% [^ 0 ^] '= 0
% [^ 1 ^] '= 1
\end{multicols}

where {{+}} is numeric addition and {{*}} is numeric multiplication.

\end{definition}
\noindent
In the setting of {{Pi}} the \emph{size}, {{[^ b ^]}}, may be simply thought of as
the number of inhabitants (an arity) of the type. This intuition will
however become tenuous in the presence of negative and fractional
types.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Graphical Language}

@@

\item Composition TODO.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Semantics}

@@

\vspace{-20pt}
    \begin{multicols}{2}
      \begin{scriptsize}

%subcode{proof} include main
%@@ f v1 |--> v2
%@ g v2 |--> v3
%@ h v3 |--> v4
%@@ g (;) h  v2 |--> v4
%@@@ f (;) (g (;) h) v1 |--> v4
~
%subcode{proof} include main
%@ f v1 |--> v2
%@ g v2 |--> v3
%@@ f (;) g v1 |--> v3
%@@ h v3 |--> v4
%@@@ (f (;) g) (;) h v1 |--> v4

      \end{scriptsize}
    \end{multicols}

@@

\begin{lemma}[Dagger]
  {{Pi}} is a dagger category, where every morphism {{f : b1<->b2}}
  has the adjoint {{f^{dagger}:b2 <->b1}}. The following properties
  hold (where {{g : b2 <-> b3}}): 
  \begin{enumerate}
  \item {{id^{dagger} = id : b <-> b}}.
  \item {{(f (;) g)^{dagger} = g^{dagger} (;) f^{dagger}: b3 <-> b1}}.
  \item {{f^{dagger dagger} = f : b1 <-> b2}}.
  \end{enumerate}
\end{lemma}

@@

  \begin{proof}
    To establish that category is monoidal one must show isomorphisms
    \begin{enumerate}
    \item {{b1 + (b2 + b3) <-> (b1 +b2) + b3}} given by {{assocl+}}.
    \item {{0 + b <-> b}} given by {{zeroe}}.
    \item {{b + 0 <-> b}} given by {{swap+ (;) zeroe}}.
    \item We need to check that {{+}} is a bifunctor. 
      \begin{enumerate}
      \item {{id_{b1<->b1}(+)id_{b2<->b2} = id_{b1+b2<->b1+b2} }}.
      \item {{(f(+)g)(;)(j(+)k) = (f (;) j) (+) (g (;) k)}}.
      \end{enumerate}
    \item We need to check the naturality of {{assocl+}}, {{zeroe}} and
      {{swap+(;)zeroe}}.
      \begin{enumerate}
      \item {{assocl+ (;) ((f (+)g) (+) h) = (f (+)(g (+) h)) (;) assocl+}}.
      \item {{zeroe (;) f = (id (+) f) (;) zeroe}}.
      \item {{(swap+ (;) zeroe) (;) f = (f (+) id) (;) (swap+ (;) zeroe)}}.
      \end{enumerate}
    \item Satisfy certain coherence conditions which are usually
      called the ``pentagon'' and ``triangle'' axioms (see Sec 3.1
      \cite{springerlink:10.1007/978-3-642-12821-94})
    \end{enumerate}

@@

    \begin{enumerate}
    \item The braiding must satisfy two ``hexagon axioms''.
    \item The braiding is self inverse,
      {{swap+_{b1+b2}=(swap+_{b2+b1})^{dagger} }}.
    \end{enumerate}
